---
title: "hw3"
author: Jiaheng Lyu
output: html_document
date: "2022-10-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(dplyr)
library(nycflights13)
library(ggplot2)
library(sjmisc)
```


## Q1 

> Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:

```{r}
airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
```

```{r}
q1 = flights %>%
  group_by(dest) %>%
  summarise(delay = mean(arr_delay, na.rm = T)) %>%
  inner_join(airports, c(dest = "faa")) %>%
  ggplot(aes(lon, lat, colour = delay)) +
  borders("state") +
  geom_point() +
  coord_quickmap()
q1
```




## Q2

> Add the location of the origin and destination (i.e. the **lat** and **lon**) to **flights**.

```{r}
q2 = flights %>%
  left_join(
    airports %>% select(faa, lat, lon),
    by = c("origin" = "faa")
  ) %>%
  left_join(
    airports %>% select(faa, lat, lon),
    by = c("dest" = "faa"),
    suffix = c("_origin", "_dest")
  )

q2[, c(1:3, 17, 13, 20, 21, 14, 22, 23)] %>%
  head(20)
```



## Q3

> Is there a relationship between the age of a plane and its delays?

+ In general, with the age of plane increasing, the delays decrease (no matter the departure delay or arrival delay) \
+ To be more detailed, delays increase with the age of plane until ten years, then it delines

```{r}
q3 = flights %>%
  inner_join(planes %>% select(tailnum, plane_year = year), by = "tailnum") %>%
  mutate(plane_age = year - plane_year) %>%
  filter(!is.na(plane_age)) %>%
  group_by(plane_age) %>%
  summarise(
    dep_delay_mean = mean(dep_delay, na.rm = TRUE),
    arr_delay_mean = mean(arr_delay, na.rm = TRUE)
  )
  
ggplot(q3, aes(x = plane_age, y = dep_delay_mean)) +
  geom_point() +
  geom_smooth(method=lm) +
  xlab("Age of plane (years)") +
  ylab("Mean Departure Delay (minutes)")

ggplot(q3, aes(x = plane_age, y = arr_delay_mean)) +
  geom_point() +
  geom_smooth(method=lm) +
  xlab("Age of Plane (years)") +
  ylab("Mean Arrival Delay (minutes)")

lm(plane_age ~ dep_delay_mean, data = q3)
lm(plane_age ~ arr_delay_mean, data = q3)
```



